On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-09-11 Revisiones/PNA_MMP_Responses to the reviewers.tex

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
%\tableofcontents

We  are grateful to the reviewers for taking their time  to review our work. Their comments have helped us to improve the paper. Changes to the original manuscript are given in  color \textcolor{red}{red} for corrections and  \textcolor{blue}{blue} for modified text in the revised version.  A detailed response to both reviewers is given below.

\section*{Response to Reviewer \#1}
{\it
The authors have presented a promissing approach for MPM to surpass the shock wave propagation in the traditional MPM. The paper was very well written and the results suggested that the proposed approach works.
}
\\
\\
The reviewer's positive commentaries are greatly appreciated.
\\

\textit{However, most of the applications presented in this manuscript are relatively simple and does not really demonstrate the capability of the proposed approach. I suggest the authors to extend their manuscript to predict more challenging problems and compare their results to the original MPM solutions.}\\

Authors agree with the proposal of the reviewer. In the first version, aiming to make a readable manuscript, several examples were not shown. In this revised version, in order to clarify several questions of both reviewers, we have introduced an additional case of study which consists in the simulation of a side impact in an elastic square plate. This test has been conveniently adapted from the work of Hammerquist \& Nairn \cite{HAMMERQUIST2017724}, where fast dynamics were also studied.

@@

\hspace{5mm}

\section*{Response to Reviewer \#2}

\textit {This is, on the face of it, an interesting piece of work but in its present form I don't think this is ready for CMAME. The paper is too long, rambling and has many grammatical and spelling errors. }\\

Authors really appreciate that the reviewer consider the present research as an interest piece of work. In order to improve the readability of the revised paper, a proofreading is made. Changes are found in the revised manuscript on \textcolor{blue}{blue} colour.\\

@@

\item \textit{Does the use of max-ent solve the problem of imposing Dirichlet boundary conditions in the MPM?}\\

Unfortunately, local \textit{max-ent} approximation technique only provides an efficient tool to transfer information from particles to nodes and vice versa to the MPM. Hence, the difficulties of imposing Dirichlet boundary conditions (DBC) in the MPM persists since they are a direct consequence of discretising the continuum with material points instead of nodes, where traditionally DBC are imposed. However, any available technique employed to overcome this situation in the standard MPM can be translated to the \textit{max-ent} one thanks to the Kronecker-Delta property inherent to it \cite{Arroyo2006}. See for instance the work of Cortis {\it et al.} (2018) \cite{Cortis_et_al_2017_IJNME} as example of how DBC can be imposed in the MPM.

\item \textit{I am sceptical of the novelty of the proposed explicit scheme as there is no mention of the recent XPIC(m) approach of Hammerquist and Nairn or similar approaches used in the Computational Graphics community, e.g. Stomakhin et al. (2013) or the more recent Polypic.}\\

These interesting publications will be added to the state of art of the revised manuscript. In the humble opinion of the authors, the XPIC approach is dedicated to mitigate the null space error produced during the transfer of information between particles and nodes. Therefore, it is related to the spatial discretisation instead of the temporal one. Meanwhile, the algorithm introduced in this research proposes the stabilisation within the adaptation of the predictor-corrector algorithm, which has been used traditionally in FEM to deal with fast dynamics. To the author\textquotesingle s knowledge, this is the first attempt to adapt the  traditional predictor-corrector scheme for the MPM framework.

@@

\end{enumerate}

 \textit{The numerical examples are not impressive when the message of the paper is that this is good for fast dynamics problems. Compare to papers such as Ma et al. (2009) IJ Impact Engineering}\\

Contrary to the reviewer\textquotesingle s opinion, the provided results are capable to highlight the strong capabilities of the MPM Newmark Predictor-Corrector scheme (NPC). It is important to remark the extension of this research: in order to isolate the strength of the proposed methodology, covering spatial and time discretization, only linear elastic problems are assessed.

Although the results presented in Ma {\it et al.} \cite{MA2009272} are impressive, the proposed examples are out of the scope of the present research. In order to reinforce the benefits of the proposed methodology, we propose a 2D wave propagation problem. It is analogous to the work of Hammerquist \& Nairn \cite{HAMMERQUIST2017724} (previously cited by the reviewer). This new test highlights the capabilities of the NPC to reproduce complex wave propagation patterns under challenging loads such as impacts. Moreover, it is capable to reduce the presence of numerical noise compared with the FE solution. Additionally, the robustness of the local \textit{max-ent} approximation dealing with unstructured mesh is depicted.

@@

\bibliographystyle{unsrt}
\begin{thebibliography}{10}

\bibitem{HAMMERQUIST2017724}
Chad C. Hammerquist and John A. Nairn.
\newblock {A new method for material point method particle updates that reduces noise and enhances stability}.
\newblock {Computer Methods in Applied Mechanics and Engineering},  318: 724--738, 2017.

\bibitem{Sadeghirad_2011}
Sadeghirad, A. and Brannon, R. M. and Burghardt, J.
\newblock {A convected particle domain interpolation technique to extend applicability of the material point method for problems involving massive deformations}.
\newblock {International Journal for Numerical Methods in Engineering}, 86(12): 1435-1456, 2011.

\bibitem{Arroyo2006}
Arroyo, M. and Ortiz, M.
\newblock{Local maximum-entropy approximation schemes: A seamless bridge between finite elements and meshfree methods}.
\newblock{International Journal for Numerical Methods in Engineering},  65(13): 2167--2202, 2006.

\bibitem{Cortis_et_al_2017_IJNME}
Cortis, Michael and Augarde, Charles and Coombs, William and Robinson, Scott and Brown, Michael and Brennan, Andrew.
\newblock {Imposition of essential boundary conditions in the material point method}.
\newblock {International Journal for Numerical Methods in Engineering}, 113: 130--152, 2017.

\bibitem{MA2009272}
Ma, S., Zhang, X.  and  Qiu, X.M.
\newblock{Comparison study of MPM and SPH in modeling hypervelocity impact problems}.
\newblock{International Journal of Impact Engineering}, 36(2): 272--282, 2009.

### 2020-09-11 LME-MPM.tex

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation in the large deformation regime is still challenging
  under this approach due \textcolor{red}{to} the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome it, we propose in this paper an explicit Predictor-Corrector time
    integration scheme. Its powerful performance mitigates the
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in
  \acrshort{mpm} occurs when the material points cross computational grid boundaries, being caused by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  surmount it, we propose in this work the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \acrshort{lme} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter,
  they are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements through numerical examples.    
\end{abstract}

@@

The aim of this document is to mitigate the spurious oscillations due to
inaccuracies in both space and time discretisation by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an explicit predictor-corrector scheme. The advantages of
this approach will be illustrated through several demanding test cases on the elastic regime: the
propagation of shock waves in an elastic bar and the response of a block of soil gradually loaded with gravitational forces.

The article is organised as follows : Section \ref{sec:meshfree-methodology}
is devoted to describe the meshfree methodology adopted in this
research, first \acrshort{mpm} procedure is introduced in
\ref{sec:derivation-mpm}, second the explicit predictor-corrector
time integration scheme is presented in \ref{sec:epc-algor-mpm}, and
third \acrshort{lme} approximation scheme will be introduced in
\ref{sec:local-max-ent}. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems}, applications to prove the numerical accuracy of the proposed approach are
presented. Finally, conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention: \textcolor{blue}{Any vectorial field is defined with an overarrow as $\vec{\square}$}. Three kind of subscripts or superscripts are used within paper. The subscript $\square_p$ is used to define a particle
variable. While the subscript $\square_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\square^{\psi}$ involves a
virtual magnitude. For the operators, the convention is: $\dot{\square}$ and
$\ddot{\square}$ for the first and second time derivative, $\square \otimes
\square$ means the dyadic operator, $\square \cdot \square$ and $\square \colon \square$ means the
first and second contraction of a tensor, $\Div{\square}$ denotes the
divergence operator, and finally $\Grad{\square}$ and $\GradS{\square}$ denotes the gradient and its symmetric part. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are \textcolor{blue}{described} in subsection \ref{sec:derivation-mpm}, \ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. We consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \Omega
= \Gamma_d \cup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \Omega$
    of the domain $\Omega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we relate global to local state by the \textit{compatibility equation} \textcolor{blue}{where the strain field \gls{strain} is defined by}:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vec{u}},
\end{equation}
\textcolor{blue}{together with essential or \acrfull{dbc} at \gls{dirichlet-boundary}.} We will further assume that 
strains are infinitesimal. The stress field \gls{stress} is considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of \textcolor{red}{linear} momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vec{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vec{b}
\end{equation}
\textcolor{blue}{together with the natural of \acrfull{nbc} at \gls{neumann-boundary}. An additional component is (iii) the constitutive equation which relates stress and strain increments as,}
{\color{red}
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\Delta \sigma} = \tens{D} \colon \tens{\Delta \varepsilon}.
\end{equation}
}
\textcolor{blue}{
In this research, plane strain linear elasticity is considered. The final equation of the set is (iv) the mass conservation, which can be expressed by} 
\begin{equation}
  \label{eq:Rho-material-derivative}
    \dot{\rho} + \rho \Div{\vec{v}} = 0.
\end{equation}

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vec{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vec{u}^{\psi} \in
  \mathcal{H}^1 \mid \vec{u}^{\psi} = \vec{0}\ \text{on}\ \Gamma_d \};
\end{equation}
\textcolor{blue}{where $\vec{u}^{\psi} \in \mathcal{H}^1$ means that}
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vec{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vec{u} \in
\mathcal{H}^1_0$ such that, for any $\vec{u}^{\psi} \in
\mathcal{H}^1_0$,
the following holds:
%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
formulation of the problem. Next, we will discretise the set of equations of the mathematical model using a double discretisation procedure, see figure~\ref{fig:MPM-discretization}.\\

\textcolor{blue}{First, the velocity field and the virtual displacements fields are discretised in a finite set of nodes $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. An approximation of the continuum field can be obtained with the help of nodal values and an appropriate interpolation functions $N_I(\vec{x})$. Also spatial derivatives of those quantities, such as gradients and divergences, are computed through the support of the background set of nodes as}

@@

\textcolor{blue}{Secondly, the continuum \gls{domain} is discretised with a finite set of material points (also denominated particles the manuscript) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, n_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$, respectively, are assigned to each material point. Furthermore, any particle field $\varphi_p$ can be approximated within the nodes in the neighbourhood of each particle, $\mathcal{B}_p \subset \mathcal{B}$, and evaluating the interpolation function in the position of each particle as}

@@

As stated previously, an explicit forward Euler algorithm has been utilized widely within the \acrshort{mpm} methodology. This scheme has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} and can be sketched by the scheme of the figure~\ref{fig:MPM_algorithm}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme_horizontal}
  }
  \caption{Description of the three steps in \acrshort{mpm} standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Other authors have proposed many others time integration alternatives
like \cite{Guilkey_2003,Charlton_2017,Tran2019e}. In the
first publication on \acrshort{mpm} \cite{Sulsky1994}, \textcolor{blue}{the nodal acceleration
was considered to update the particles by}
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ N_{Ip}\ \vec{a}_{I}^{k},
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ N_{Ip}\ \vec{v}_{I}^{k}.
\end{equation}
However, as Andersen (2009)\cite{thesis_Andersen_2009} pointed out, this algorithm has been shown to be numerically unstable due to \textcolor{red}{the fact} that nodal forces can be infinite for an infinitesimal nodal mass $\tens{m}_I$. This issue may lead to numerical problems when nodal acceleration is obtained in the evaluation of the Eqs. \eqref{eq:Sulsky-1994-UL-x} and \eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm was proposed by Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}:
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ \frac{N_{Ip}\ \vec{p}_{I}^{k}}{\tens{m}_I}, 
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ \frac{N_{Ip}\ \vec{f}_{I}^{k}}{\tens{m}_I}.
\end{equation}
Delving into the improvement of the accuracy of the \acrshort{mpm} explicit schemes, Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for \acrshort{mpm} inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both \textcolor{red}{at} the beginning and \textcolor{red}{at} the end of the time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\  N_{Ip}\ \left[(1 - \gamma)\ \vec{a}_I^{k} +
    \gamma\ \vec{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} N_{Ip} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right],
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vec{a}_p^{k+1} = \sum_{I\ \in\ \mathcal{B}_p} N_{Ip}\ \vec{a}_{I}^{k+1}.
\end{equation}

@@

where $p_I$ \textcolor{blue}{stands for probability of the random variable outcomes. By interpreting these probabilities as the shape functions $N_I(\vec{x})$ of an approximation scheme, equation \eqref{eq:Shannon-entropy} can be regarded as a measure of the uncertainty of the approximation. Thus, according to Jaynes's principle of maximum entropy \cite{Jaynes1957},} the least-biased approximation scheme is
given by

>>>>

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an alternative to \acrfull{fem} under large
  \textcolor{blue}{deformations}. However, the simulation of shock waves
  propagation in the large deformation regime is still challenging  under this approach due to the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome it, we propose in this paper an explicit Predictor-Corrector time
    integration scheme. Its powerful performance mitigates the
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in
  \acrshort{mpm} occurs when the material points cross computational grid boundaries, being caused by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  surmount it, we propose in this work the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \textcolor{blue}{The performance of both improvements is validated by the high quality
   results of the numerical examples}.
\end{abstract}

@@

The aim of this document is to mitigate the spurious oscillations due to
inaccuracies in both space and time discretisation by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of a predictor-corrector scheme. The advantages of
this approach will be illustrated through several demanding test cases on the elastic regime: the
propagation of shock waves in an elastic bar and the response of a block of soil gradually loaded with gravitational forces.

The article is organised as follows : Section \ref{sec:meshfree-methodology}
is devoted to describe the meshfree methodology adopted in this
research, first \acrshort{mpm} procedure is introduced in
\ref{sec:derivation-mpm}, second the predictor-corrector
time integration scheme is presented in \ref{sec:epc-algor-mpm}, and
third \acrshort{lme} approximation scheme will be introduced in
\ref{sec:local-max-ent}. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems}, applications to prove the numerical accuracy of the proposed approach are
presented. Finally, conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. \textcolor{blue}{The described methodology is sketched by the scheme of the figure~\ref{fig:MPM_algorithm}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme_horizontal}
  }
  \caption{Description of the three steps in \acrshort{mpm} standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In what follows, we will adopt the following convention: \textcolor{blue}{Any vector field is defined with an overarrow as $\vec{\square}$}. Three kind of subscripts or superscripts are  considered:  subscript $\square_p$ is used to define a particle variable,  subscript $\square_I$ is reserved for nodal variables, superscript $\square^{\psi}$ involves a virtual magnitude. \textcolor{blue}{The convention adopted for the operators is as folows:} $\dot{\square}$ and $\ddot{\square}$ \textcolor{blue}{are considered} for the first and second time derivative, $\square \otimes \square$ means the dyadic operator, $\square \cdot \square$ and $\square \colon \square$ \textcolor{blue}{are the single and double contraction of tensor index}, $\Div{\square}$ denotes the divergence operator, and finally $\Grad{\square}$ and $\GradS{\square}$ denotes the gradient and its symmetric part. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are \textcolor{red}{described} in subsection \ref{sec:derivation-mpm}, \ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. We consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \Omega
= \Gamma_d \cup \Gamma_n$  \textcolor{red}{where} $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \Omega$
    of the domain $\Omega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we relate global to local state by the \textit{compatibility equation} \textcolor{blue}{where the strain field \gls{strain} is defined by}:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vec{u}},
\end{equation}
\textcolor{blue}{together with essential or \acrfull{dbc} at \gls{dirichlet-boundary}.} We will further assume that 
strains are infinitesimal. The stress field \gls{stress} is considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of \textcolor{red}{linear} momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vec{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vec{b}
\end{equation}
\textcolor{blue}{together with the natural of \acrfull{nbc} at \gls{neumann-boundary}. An additional component is (iii) the constitutive equation which relates stress and strain increments as,}
{\color{red}
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\Delta \sigma} = \tens{D} \colon \tens{\Delta \varepsilon}.
\end{equation}
}
\textcolor{blue}{
In this research, plane strain linear elasticity is considered. The final equation of the set is (iv) the mass conservation, which can be expressed by} 
\begin{equation}
  \label{eq:Rho-material-derivative}
    \dot{\rho} + \rho \Div{\vec{v}} = 0.
\end{equation}

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vec{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vec{u}^{\psi} \in
  \mathcal{H}^1 \mid \vec{u}^{\psi} = \vec{0}\ \text{on}\ \Gamma_d \};
\end{equation}
\textcolor{blue}{where $\vec{u}^{\psi} \in \mathcal{H}^1$ means that}
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vec{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vec{u} \in
\mathcal{H}^1_0$ such that, for any $\vec{u}^{\psi} \in
\mathcal{H}^1_0$,
the following holds:
%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
\sout{formulation} of the problem. Next, we will discretise the set of equations of the mathematical model using a double discretisation procedure, see figure~\ref{fig:MPM-discretization}.\\

\textcolor{blue}{First, the velocity field and the virtual displacements fields are discretised in a finite set of nodes $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. An approximation of the continuum field can be obtained with the help of nodal values and \sout{an} appropriate interpolation functions $N_I(\vec{x})$. Also spatial derivatives of those quantities, such as gradients and divergences, are computed through the support of the background set of nodes as}

@@

\textcolor{blue}{Secondly, the continuum \gls{domain} is discretised with a finite set of material points (also known as particles in this manuscript) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, n_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$, respectively, are assigned to each material point. Furthermore, any particle field $\varphi_p$ can be approximated within the nodes in the neighbourhood of each particle, $\mathcal{B}_p \subset \mathcal{B}$, and evaluating the interpolation function in the position of each particle as}

@@

As stated previously, an explicit forward Euler algorithm has been utilized widely within the \acrshort{mpm} methodology. This scheme has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009}. Other authors have proposed many others time integration alternatives like \cite{Guilkey_2003,Charlton_2017,Tran2019e}. In the
first publication on \acrshort{mpm} \cite{Sulsky1994}, \textcolor{blue}{the nodal acceleration
was considered to update the particles by}
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ N_{Ip}\ \vec{a}_{I}^{k},
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ N_{Ip}\ \vec{v}_{I}^{k}.
\end{equation}
However, as Andersen (2009)\cite{thesis_Andersen_2009} pointed out, this algorithm has been shown to be numerically unstable due to \textcolor{red}{the fact} that nodal forces can be infinite for an infinitesimal nodal mass $\tens{m}_I$. This issue may lead to numerical problems when nodal acceleration is obtained in the evaluation of the Eqs. \eqref{eq:Sulsky-1994-UL-x} and \eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm was proposed by Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}:
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ \frac{N_{Ip}\ \vec{p}_{I}^{k}}{\tens{m}_I}, 
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ \frac{N_{Ip}\ \vec{f}_{I}^{k}}{\tens{m}_I}.
\end{equation}
Delving into the improvement of the accuracy of the \acrshort{mpm} explicit schemes, Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for \acrshort{mpm} inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both \textcolor{red}{at} the beginning and \textcolor{red}{at} the end of the time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\  N_{Ip}\ \left[(1 - \gamma)\ \vec{a}_I^{k} +
    \gamma\ \vec{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} N_{Ip} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right],
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vec{a}_p^{k+1} = \sum_{I\ \in\ \mathcal{B}_p} N_{Ip}\ \vec{a}_{I}^{k+1}.
\end{equation}

@@

where $p_I$ \textcolor{blue}{stands for probability of the random variable outcomes. By interpreting these probabilities as the shape functions $N_I(\vec{x})$ of an approximation scheme, equation \eqref{eq:Shannon-entropy} can be regarded as a measure of the uncertainty of the approximation. Thus, according to Jaynes's principle of maximum entropy \cite{Jaynes1957},} the least-biased approximation scheme  \textcolor{red}{can be} given by
